---
title: "The Djesse Albums"
author: "Lucas Straathof"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical layout: scroll
    storyboard: true 
    theme: journal
---

```{r, include=FALSE}
library(tidyverse)
library(spotifyr)
library(ggplot2)
library(plotly)
library(dplyr)
```

```{r, include=FALSE}
djesse1 <- get_playlist_audio_features("", "55AhOVOznsMmppxhkhQeox")
djesse2 <- get_playlist_audio_features("", "2ySg6xH1guPJ6WacYBnEtE")
djesse3 <- get_playlist_audio_features("", "4y0xPsPUbwmsf32r8GiETo")
all_albums <- bind_rows(djesse1, djesse2, djesse3)
```


### Home - The Djesse Albums (This is going to be the home page where I give an introduction and tell the people my corpus)

Welcome to my page about the wonderful albums of Jacob Collier, called Djesse. This series of albums covers tons of genres and are completely different from each other. In this research, I'm going to dive deeper into these albums, seeing if there are comparisons to be made, or to see if they are completely different. 

### The Albums (This page is going to be filled with information about the albums themselves and what their about)




### First Comparisons (On this page, I'm going to give a comparison of the albums)

```{r, include=FALSE}
energy_of_all <- ggplot(all_albums, aes(x=energy, fill=playlist_name)) +
  geom_histogram(binwidth =0.05) +
  theme_classic()
```

```{r}
ggplotly(energy_of_all)
```

***

In this first graph, we compare the energy of all the albums. 

