---
title: "The Djesse Albums"
author: "Lucas Straathof"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical layout: scroll
    storyboard: true 
    theme: journal
---

```{r, include=FALSE}
library(tidyverse)
library(tidymodels)
library(compmus)
library(protoclust)
library(cowplot)
library(spotifyr)
library(ggplot2)
library(plotly)
library(dplyr)
library(ggdendro)
library(heatmaply)

get_conf_mat <- function(fit) {outcome <- .get_tune_outcome_names(fit)
  fit |> 
    collect_predictions() |> 
    conf_mat(truth = outcome, estimate = .pred_class)
}  

get_pr <- function(fit) {
  fit |> 
    conf_mat_resampled() |> 
    group_by(Prediction) |> mutate(precision = Freq / sum(Freq)) |> 
    group_by(Truth) |> mutate(recall = Freq / sum(Freq)) |> 
    ungroup() |> filter(Prediction == Truth) |> 
    select(class = Prediction, precision, recall)
}  

```

```{r, include=FALSE}
djesse1 <- get_playlist_audio_features("", "55AhOVOznsMmppxhkhQeox")
djesse2 <- get_playlist_audio_features("", "2ySg6xH1guPJ6WacYBnEtE")
djesse3 <- get_playlist_audio_features("", "4y0xPsPUbwmsf32r8GiETo")
djesse4 <- get_playlist_audio_features("", "23CJn0kvWIvoEg6VlbGeGl")

all_albums <- bind_rows(djesse1, djesse2, djesse3, djesse4)
```



### Home - The Djesse Albums

Welcome to my page about the wonderful albums of Jacob Collier, called Djesse. This series of albums covers tons of genres and are completely different from each other. In this research, I'm going to dive deeper into these albums, seeing if there are comparisons to be made, or to see if they are completely different. 

There is a fourth Djesse album on its way and will be released at the end of February, but I think it’s easier to analyze if I can start right away, seeing as we're still in the middle of February as of the point we started this course. I might change this if it becomes clear that I can easily include it into my corpus

I really like Jacob Collier’s music and am a big fan. His Djesse albums especially blow me out of the water each time. His music is so different from anyone else’s. He has his very distinct style which is instantly recognizable. Jacob Collier is his own genre within music in a way. 

The comparisons will be within the albums themselves. I find it interesting to discover if there is a difference between the Djesse albums in a way I haven't discovered before. I’m expecting that there will be a constant theme within a single album, but differences between the few. 

The thing that I think is going to be the most troubling, is the fact that there are not that many songs,  around 50 actually. This will become an issue if it shows that I need more data to do the research. For example, if I’m seeing a correlation, but that correlation is dependent on 2 or 3 songs being the same in a way, that isn’t much evidence. With more songs, this evidence would be stronger. 

I think the biggest outlying song will be ‘Moon River’ on Djesse vol. 2. This song was completely sung acapella and is very different from everything else on the albums, even though every single thing Jacob Collier makes is varying in every aspect.


This is a test



