---
title: "The Djesse Albums"
author: "Lucas Straathof"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical layout: scroll
    storyboard: true 
    theme: journal
---

```{r, include=FALSE}
library(tidyverse)
library(tidymodels)
library(compmus)
library(protoclust)
library(cowplot)
library(spotifyr)
library(ggplot2)
library(plotly)
library(dplyr)
library(ggdendro)
library(heatmaply)

get_conf_mat <- function(fit) {outcome <- .get_tune_outcome_names(fit)
  fit |> 
    collect_predictions() |> 
    conf_mat(truth = outcome, estimate = .pred_class)
}  

get_pr <- function(fit) {
  fit |> 
    conf_mat_resampled() |> 
    group_by(Prediction) |> mutate(precision = Freq / sum(Freq)) |> 
    group_by(Truth) |> mutate(recall = Freq / sum(Freq)) |> 
    ungroup() |> filter(Prediction == Truth) |> 
    select(class = Prediction, precision, recall)
}  

```

```{r, include=FALSE}
djesse1 <- get_playlist_audio_features("", "55AhOVOznsMmppxhkhQeox")
djesse2 <- get_playlist_audio_features("", "2ySg6xH1guPJ6WacYBnEtE")
djesse3 <- get_playlist_audio_features("", "4y0xPsPUbwmsf32r8GiETo")
djesse4 <- get_playlist_audio_features("", "23CJn0kvWIvoEg6VlbGeGl")
djesse3_4 <- bind_rows(djesse3, djesse4) |> 
  add_audio_analysis()
all_albums <- bind_rows(djesse1, djesse2, djesse3, djesse4)
```


### Home - The Djesse Albums

Welcome to my page about the wonderful albums of Jacob Collier, called Djesse. This series of albums covers tons of genres and are completely different from each other. In this research, I'm going to dive deeper into these albums, seeing if there are comparisons to be made, or to see if they are completely different. 

There is a fourth Djesse album on its way and will be released at the end of February, but I think it’s easier to analyze if I can start right away, seeing as we're still in the middle of February as of the point we started this course. I might change this if it becomes clear that I can easily include it into my corpus

I really like Jacob Collier’s music and am a big fan. His Djesse albums especially blow me out of the water each time. His music is so different from anyone else’s. He has his very distinct style which is instantly recognizable. Jacob Collier is his own genre within music in a way. 

The comparisons will be within the albums themselves. I find it interesting to discover if there is a difference between the Djesse albums in a way I haven't discovered before. I’m expecting that there will be a constant theme within a single album, but differences between the few. 

The thing that I think is going to be the most troubling, is the fact that there are not that many songs,  around 50 actually. This will become an issue if it shows that I need more data to do the research. For example, if I’m seeing a correlation, but that correlation is dependent on 2 or 3 songs being the same in a way, that isn’t much evidence. With more songs, this evidence would be stronger. 

I think the biggest outlying song will be ‘Moon River’ on Djesse vol. 2. This song was completely sung acapella and is very different from everything else on the albums, even though every single thing Jacob Collier makes is varying in every aspect.

### The Albums

In 2018, Collier began working on Djesse, a four-volume, 50-song album featuring more than two dozen artists and ensembles. The first volume, which features the Metropole Orkest, Djesse Vol. 1, was released in December 2018. The second, Djesse Vol. 2, uses more acoustic instrumentation, and was released in July 2019. In 2020, Collier won Grammy Awards for his arrangements of "All Night Long (All Night)" from Djesse Vol. 1 and "Moon River" from Djesse Vol. 2. In 2021, he won a Grammy Award for "He Won't Hold You" from the third volume, Djesse Vol. 3. 

[Wikipedia Page](https://en.wikipedia.org/wiki/Jacob_Collier)


### First Comparison

```{r, include=FALSE}
energy_of_all <- ggplot(all_albums, aes(x=energy, fill=playlist_name, alpha = 0.3)) +
  geom_density() +
  scale_x_continuous(breaks = seq(0, 1, by = 0.1)) +
  facet_wrap(~ factor(playlist_name, levels = c('Djesse Vol. 1', 'Djesse Vol. 2', 'Djesse Vol. 3', 'Djesse Vol. 4'))) +
  theme_classic()
```

```{r}
ggplotly(energy_of_all)
```

***

In this first graph, we compare the energy of all the albums. And we can see a few differences. 
First of all, Djesse Vol. 3 seems to be the most energetic overall, with nothing going beneath the 0.3 energy range. 
Secondly, Djesse Vol. 2 is very spread out. There are songs on all parts of the graph.


### Dendogram

```{r}
djesse_3_4_dendogram <-
  get_playlist_audio_features("", "4LGploi0y23ssKcueAa98J") |>
  add_audio_analysis() |> 
  mutate(segments = map2(segments, key, compmus_c_transpose), 
         pitches = map(segments, compmus_summarise, pitches, method = "mean", norm = "manhattan"),
         timbre = map(segments, compmus_summarise, timbre, method = "mean")) |>
  mutate(pitches = map(pitches, compmus_normalise, "clr")) |>
  mutate_at(vars(pitches, timbre), map, bind_rows) |>
  unnest(cols = c(pitches, timbre))
```

```{r}
replace_track_names <- function(track_name) {
  track_name <- gsub("Pt\\. 1", "Part 1", track_name)
  track_name <- gsub("Pt\\. 2", "Part 2", track_name)
  track_name <- gsub("^(.*?) (.*)$", "\\2 \\1", track_name)
  return(track_name)}

djesse3_4$track.name <- sapply(djesse3_4$track.name, replace_track_names)
```


```{r}
djesse_3_4_dendogram_computed <-
  recipe(
    track.name ~
      danceability + energy + loudness + speechiness + acousticness + 
      instrumentalness + liveness + valence + tempo,
    data = djesse3_4
  ) |>
  step_center(all_predictors()) |>
  step_scale(all_predictors()) |> 
  # step_range(all_predictors()) |> 
  prep(djesse3_4 |> mutate(track.name = str_trunc(track.name, 20))) |>
  juice() |>
  column_to_rownames("track.name")

djesse_3_4_dendogram_computed <- dist(djesse_3_4_dendogram_computed, method = "euclidean")

data_for_djesse_clustering <- djesse_3_4_dendogram_computed |> 
  hclust(method = "average") |> # average for a balanced tree!
  dendro_data() 

playlist_data_for_join <- djesse3_4 %>%
  select(track.name, playlist_name) %>%
  mutate(label = str_trunc(track.name, 20))

data_for_djesse_clustering$labels <- data_for_djesse_clustering$labels %>%
  left_join(playlist_data_for_join)

data_for_djesse_clustering$labels$label <- factor(data_for_djesse_clustering$labels$label)
```

```{r}
plot_dendogram <- data_for_djesse_clustering |>
  ggdendrogram() +
  geom_text(data = label(data_for_djesse_clustering), aes(x, y, 
                                   label=label, 
                                   hjust=0, 
                                   colour=playlist_name), size=3) +
  coord_flip() + 
  scale_y_reverse(expand=c(0.2, 0)) +
  theme(axis.line.y=element_blank(),
        axis.ticks.y=element_blank(),
        axis.text.y=element_blank(),
        axis.title.y=element_blank(),
        panel.background=element_rect(fill="white"),
        panel.grid=element_blank()) +
  labs(title = "Playlist Clustering for Djesse Vol. 3 and 4") +
  guides(
    colour = guide_legend(
      title = "Playlist"
    )
  )
```

```{r, fig.width=10, fig.height=6}
plot_dendogram
```


***

This is some text as a test.



### Key Centers

```{r, include=FALSE}
all_albums$mode <- as.factor(all_albums$mode)

key_labels_x <- c("C", "C#", "D", "Eb", "E", "F", "F#", "G", "Ab", "A", "Bb", "B")
mode_labels_y <- c("minor", "major")
mode_colors_y <- c("red", "blue")

theme_key_center <- theme(axis.text.x = element_text(hjust = 1),
        axis.title.y = element_blank(), 
        panel.grid.major = element_blank(),  
        panel.background = element_rect(fill = "white"),  
        panel.grid.minor = element_line(color = "gray"),
        plot.title = element_text(hjust = 0.5, margin = margin(b = 20))
)

key_center <- ggplot(all_albums, aes(x=key, fill=mode)) +
              geom_histogram(binwidth = 1, color = "gray25") +
              facet_wrap(~ factor(playlist_name, levels = c('Djesse Vol. 1', 'Djesse Vol. 2', 'Djesse Vol. 3', 'Djesse Vol. 4'))) +
              scale_x_continuous(breaks = 0:11, labels = key_labels_x) +
              scale_y_continuous(limits = c(0, 4)) +
              scale_fill_manual(labels = mode_labels_y, values = mode_colors_y) + 
              expand_limits(x = c(0, 11)) +
              ggtitle("Djesse Albums Key Centers") +
              theme_key_center
```


```{r, fig.width=10, fig.height=6}
ggplotly(key_center)
```

***

In these graphs, we see the key centers of the three different albums. First of all, the Djesse Vol. 1 album purely contains major songs, which is interesting. Actually, over all the songs there aren't that many minor songs. So I'd say this is a big correlation between them all.

It's also funny to see that Collier has never put a song in the key of Eb or G into a Djesse album. 


### Tempo, Volume and Duration

```{r}
djesse_vol <- all_albums |> add_audio_analysis()
```

```{r}
djesse_vol <- djesse_vol |> 
  mutate(sections =
      map(sections, summarise_at, 
      vars(tempo, loudness, duration),
      list(section_mean = mean, section_sd = sd)
      )) |>
  unnest(sections) |> 
  ggplot(aes(x = tempo, y = tempo_section_sd, color = playlist_name, alpha = loudness)) +
    geom_point(aes(size = duration / 60)) +
    geom_rug() +
    facet_wrap(~ factor(playlist_name, levels = c('Djesse Vol. 1', 'Djesse Vol. 2', 'Djesse Vol. 3', 'Djesse Vol. 4'))) +
    theme_minimal() +
    ylim(0, 5) +
    labs(x = "Mean Tempo (bpm)", y = "SD Tempo", color = "Album name", size = "Duration (min)", alpha = "Volume (dBFS)", 
         title = "Tempo, Volume and Duration")
```

```{r, fig.width=10, fig.height=6}
djesse_vol
```

***

In these graphs, it shows the tempo's, volumes and the duration of the songs of all four albums and we can see some interesting things. 

In the first album, we can see that there is a cluster of songs which have relatively the same mean and SD tempo, with only one outlier. 

In the second album, there seem to be 3 different columns of mean tempo, the first around 90 bpm, the second around 120 bpm and the last around 145 bpm. I don't think this is a real correlation though. I'd doubt if Jacob did this on purpose. The more logical reason for this, is that these are standard tempo's for songs which aren't meant to be really fast or really slow. 
Although, it is interesting to see that the songs here deviate a lot from the mean, at least, more than the other albums. If I were to guess what the biggest outlier is (with a standard deviation of around 4.1), I'd guess it is the song '[Feel](https://open.spotify.com/track/7MGNHuYwmm9UjQgdVciO1v?si=a32fe2cb1042488f)'. This song, as the title suggests, has a great feel to it, which makes the rhythm sway more than usual. 

The third album is more consistent with it's deviation in tempo, but is more random with it's mean tempo, having the biggest differences in this category. 

The thing I'm most troubled by, is the fact that in the fourth album, I don't see a big outlier. [Box of Stars Pt. 2](https://open.spotify.com/track/232QdHfiXCR7dqhScytiYN?si=6914eff7871846a2) is a very interesting song and there's a moment where the rhythm gradually comes to a complete stop, slowing down over a longer period of time. It's shocking to not see this in the graph. 


### Tempogram - [Box of Stars Pt. 2]

```{r}
box_stars <- get_tidy_audio_analysis("232QdHfiXCR7dqhScytiYN")
```

```{r}
box_stars <- box_stars |>
  tempogram(window_size = 8, hop_size = 1, cyclic = TRUE) |>
  ggplot(aes(x = time, y = bpm, fill = power)) +
  geom_raster() +
  scale_fill_viridis_c(guide = "none") +
  labs(x = "Time (s)", y = "Tempo (BPM)", title = "Box of Stars Pt. 2") +
  theme_classic()
```

```{r, fig.width=10, fig.height=6}
box_stars
```

***

In this tempogram, I have plotted a tempogram of the song [Box of Stars Pt. 2](https://open.spotify.com/track/232QdHfiXCR7dqhScytiYN?si=6914eff7871846a2) from the fourth Djesse album. On the previous storyboard, I mentioned it was strange to not see this song stick out from the rest in the standard deviation, because the tempo deviates a lot. In this tempogram, you can really see this in this plot. The first few minutes are really steady in tempo, but around 90 s, the tempo keeps on declining, all the way till 160 s, where it starts to climb back up again, until 205 s, when it arrives at its original tempo. 

The next notable part is from 250 s onwards. In this section, a choir takes over and sings completely rubato. Sometimes an electric guitar jumps in with some riffs, of which it's uncertain if it's actually following a similar rhythm as the choir. These guitar riffs are also really fast, which makes it hard to get a good tempo feeling going. This can also be seen in the graph. Sometimes there is a small yellow line at random bpm amounts. I believe these are the points the e-guitar jumps in. 


### Chroma Features - [Moon River]

```{r}
moonriver_chrome <- get_tidy_audio_analysis("1IRRqn75jaZHp7zMT1NGbZ") |>
  select(segments) |>
  unnest(segments) |>
  select(start, duration, pitches)

moonriver_chrome_plot <- moonriver_chrome |>
  mutate(pitches = map(pitches, compmus_normalise, "euclidean")) |>
  compmus_gather_chroma() |> 
  ggplot(aes(x = start + duration / 2, width = duration, y = pitch_class, fill = value)) +
  geom_tile() +
  geom_vline(xintercept = 228, color = "red", size = 1.35) +
  geom_vline(xintercept = 274, color = "red", size = 1.35) +
  geom_vline(xintercept = 347, color = "red", size = 1.35) +
  geom_vline(xintercept = 410, color = "red", size = 1.35) +
  geom_vline(xintercept = 474, color = "red", size = 1.35) +
  labs(x = "Time (s)", y = NULL, fill = "Magnitude", title = "Moonriver (Jacob Collier)") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, margin = margin(b = 20, t = 20))) +
  scale_fill_viridis_c()
```

```{r, fig.width=10, fig.height=5}
plot_grid(moonriver_chrome_plot, ncol = 1)
```

***

In this Chromagram plot, I plotted the song Moon River by Jacob Collier. That is because he transposes this song tons and he even goes to keys which don't exist on a normal piano, like half sharps. That is why I thought it would be interesting to see what the chroma plot would do. 
The transcription which I'm going to be following is on YouTube and the video is [here](https://www.youtube.com/watch?v=KhyyjRcrn84)

The song starts in Bb major, then goes to Db major around 3 minutes and 48 seconds (228 s) and then around 4.34 (247 s) he transposes to D Major, up a semitone. At around 5.47 (347 s), Jacob goes to D Half Sharp Major, which is up a quarter tone. After that, he goes to another weird key. At 6.50 (410 s), he goes to E Half Flat Major. Then, he does something even more interesting. He gradually rises the key over a longer period of time, from 7.54 (474 s) to 8.18 (which is right at the end of the song) it transposes from E Half Flat to E Half Sharp. 

I think the most obvious point where you can see that Jacob is outside the 12 tone temperament is between 410 s and 474 s, where the song is in E Half Flat Major. This is the point where you can see the most dark spaces, which means that the system doesn't really know which notes are being played. 



### Chordogram - 'Moonriver'

```{r}
circshift <- function(v, n) {
  if (n == 0) v else c(tail(v, n), head(v, -n))
}

#      C     C#    D     Eb    E     F     F#    G     Ab    A     Bb    B
major_chord <-
  c(   1,    0,    0,    0,    1,    0,    0,    1,    0,    0,    0,    0)
minor_chord <-
  c(   1,    0,    0,    1,    0,    0,    0,    1,    0,    0,    0,    0)
seventh_chord <-
  c(   1,    0,    0,    0,    1,    0,    0,    1,    0,    0,    1,    0)

major_key <-
  c(6.35, 2.23, 3.48, 2.33, 4.38, 4.09, 2.52, 5.19, 2.39, 3.66, 2.29, 2.88)
minor_key <-
  c(6.33, 2.68, 3.52, 5.38, 2.60, 3.53, 2.54, 4.75, 3.98, 2.69, 3.34, 3.17)

chord_templates <-
  tribble(
    ~name, ~template,
    "C:maj", circshift(major_chord, 0), "C:min", circshift(minor_chord, 0), "C:7", circshift(seventh_chord, 0),
    "Db:maj", circshift(major_chord, 1), "Db:min", circshift(minor_chord, 1), "Db:7", circshift(seventh_chord, 1),
    "D:maj", circshift(major_chord, 2), "D:min", circshift(minor_chord, 2), "D:7", circshift(seventh_chord, 2),
    "Eb:maj", circshift(major_chord, 3), "Eb:min", circshift(minor_chord, 3), "Eb:7", circshift(seventh_chord, 3),
    "E:maj", circshift(major_chord, 4), "E:min", circshift(minor_chord, 4), "E:7", circshift(seventh_chord, 4),
    "F:maj", circshift(major_chord, 5), "F:min", circshift(minor_chord, 5), "F:7", circshift(seventh_chord, 5),
    "F#:maj", circshift(major_chord, 6), "F#:min", circshift(minor_chord, 6), "F#:7", circshift(seventh_chord, 6),
    "G:maj", circshift(major_chord, 7), "G:min", circshift(minor_chord, 7), "G:7", circshift(seventh_chord, 7),
    "Ab:maj", circshift(major_chord, 8), "Ab:min", circshift(minor_chord, 8), "Ab:7", circshift(seventh_chord, 8),
    "A:maj", circshift(major_chord, 9), "A:min", circshift(minor_chord, 9), "A:7", circshift(seventh_chord, 9),
    "Bb:maj", circshift(major_chord, 10), "Bb:min", circshift(minor_chord, 10), "Bb:7", circshift(seventh_chord, 10),
    "B:maj", circshift(major_chord, 11), "B:min", circshift(minor_chord, 11), "B:7", circshift(seventh_chord, 11)
  )
```

```{r}
moonriver_key_est <- get_tidy_audio_analysis("1IRRqn75jaZHp7zMT1NGbZ") |>
  compmus_align(bars, segments) |>
  select(bars) |>
  unnest(bars) |>
  mutate(pitches = map(segments, compmus_summarise, pitches, method = "mean", norm = "euclidean"))
```

```{r}
moonriver_key_est_plot <- moonriver_key_est |> 
  compmus_match_pitch_template(chord_templates, method = "euclidean", norm = "manhattan") |>
  ggplot(aes(x = start + duration / 2, width = duration, y = name, fill = 1 - d)) +
  geom_tile() +
  scale_fill_viridis_c() +
  theme_minimal() +
  labs(x = "Time (s)", y = "", fill = "Magnitude", title = "Moonriver (Jacob Collier)") +
  theme(plot.title = element_text(hjust = 0.5, margin = margin(b = 20, t = 20))) 
```

```{r, fig.width=10, fig.height=5}
plot_grid(moonriver_key_est_plot, ncol = 1)
```

Some small notes:

On this page, I have created a chordograph from the same song which I used for the spectogram, which is [Moonriver](https://www.youtube.com/watch?v=KhyyjRcrn84). For this chordograph, I have put the notes going down chromatically, instead of having them align with the circle of fiths. I've also made it so the yellow spots are the points that most align with what's going on in the song. 

***

Now that that's been said, let's analyse the graph. First of all, we can see tons of yellow and green spaces. This makes sense, because in this song, Jacob managed to squeeze in a lot of different chords, most of the times having more than one chord in a single bar. The song starts in Bb and we can also see this in the graph. There is an entire bar of Bb chords up until 215 s. 
In the Bb Major section, we can also see big blue lines running through the whole graph vertically. The biggest is at about 160 s. At this point in time, there are more melodies coming in, instead of it only being chords.

At 215, there seems to be a strange line going through the whole graph, which is where Jacob starts the transition to Db Major. In this section, he uses chords with notes that are very close together which change every 8th note, so a lot of chords come by in a small amount of time. After this small bar, the Db chords of course become more yellow in the graph, especially Db7. 

As to why the seventh chords are more yellow overall in this graph. I think I have an explanation. In this song, Jacob uses all kinds of chords with upper extensions, such as 7ths, 9ths, 11ths, etc... I believe this is also the reason why, throughout the whole graph, there is more than one chord possible as the correct chord within this system with basic chords

The last point I want to talk about, is the very thin line at about 395 s. What happens here, is that the base note is falling very fast for over an octave, which creates a crazy effect. This isn't a sound which is regularly heard and I'm guessing the system also doesn't really know what chord to make of it. 



### Cepstogram - 'Sleeping on my Dreams'

```{r}
sleeping_dreams <-
  get_tidy_audio_analysis("4mTd21pyWajN252LcouVIm") |>
  compmus_align(bars, segments) |>
  select(bars) |>
  unnest(bars) |>
  mutate(pitches = map(segments, compmus_summarise, pitches, method = "rms", norm = "manhattan")) |>
  mutate(timbre = map(segments, compmus_summarise, timbre, method = "rms", norm = "manhattan"))
```

```{r}
sleeping_dreams_cepstogram <- sleeping_dreams |>
  compmus_gather_timbre() |>
  ggplot(aes(x = start + duration / 2, width = duration, y = basis, fill = value)) +
  geom_tile() +
  labs(x = "Time (s)", y = NULL, fill = "Magnitude") +
  scale_fill_viridis_c() +                              
  theme_classic()
```

```{r, fig.width=10, fig.height=5}
sleeping_dreams_cepstogram
```


### SSM Graphs - 'Sleeping on my Dreams'


```{r, include=FALSE}
sleeping_dreams_pitches <- sleeping_dreams |>
  compmus_self_similarity(pitches, "euclidean") |> 
  ggplot(aes(x = xstart + xduration / 2, width = xduration, y = ystart + yduration / 2, height = yduration, fill = d)) +
  geom_tile() +
  coord_fixed() +
  scale_fill_viridis_c(guide = "none") +
  theme_classic() +
  labs(x="", y="")

sleeping_dreams_timbre <- sleeping_dreams |>
  compmus_self_similarity(timbre, "euclidean") |> 
  ggplot(aes(x = xstart + xduration / 2, width = xduration, y = ystart + yduration / 2, height = yduration, fill = d)) +
  geom_tile() +
  coord_fixed() +
  scale_fill_viridis_c(guide = "none") +
  theme_classic() +
  labs(x="", y="")

```

```{r, fig.width=10, fig.height=5.5}
plot_grid(sleeping_dreams_pitches, sleeping_dreams_timbre, ncol=2, labels = c("Pitch", "Timbre"))
```

***

In this cepstogram, we can see the pitch and timbre of the song 'Sleeping on my Dreams' by Jacob Collier from his 3rd Djesse album. I chose this song, because it has a more usual structure, with clear verses and chorus'. And we can also see this in the graphs. In the pitch graph (and also in the timbre graph), we can see clear lines where a new section begins. It first begins with a small intro, then goes to a verse, which is the first square. Then a chorus starts which is the second square. The chorus also has some different instruments, which is why you can also see this change in the timbre graph. After the chorus, another verse and chorus follow. 

In the bridge section, we can see that it changes entirely. This is because most instruments cut out at this point and it changes to a more flowing section with smooth sounds, contrasting with the more catchy and staccato instruments from the verses and chorus'. The outro is also made up out of these flowing instruments, which is why the outro and bridge look the same in the graphs. 


